A new form of transcorrelated Hamiltonian inspired by range-separated DFT

### 2020-11-12 cusp_j/paper/paper.tex
restructuration

\definecolor{darkgreen}{HTML}{009900}
\usepackage[normalem]{ulem}
\newcommand{\sphi}[1]{\hat{{\bf S}}_{#1}}
\newcommand{\overlap}[2]{\langle #1 | #2 \rangle}
\newcommand{\matelem}[3]{\langle #1 | #2 | #3 \rangle}
\newcommand{\deriv}[3]{\frac{\partial^{#3} #1}{\partial {#2}^{#3}}}
\newcommand{\bd}[1]{{\bf {#1}}}
\newcommand{\br}[0]{{\bf {r}}}
\newcommand{\bs}[0]{{\bf {s}}}
\newcommand{\dr}[1]{\text{d}{\bf {#1}}}
\newcommand{\psiex}[0]{\Psi^{\text{ex}}_0}
\newcommand{\phiex}[0]{\Phi^{\text{ex}}_0}
\newcommand{\phimu}[0]{\Phi^{\text{ex},\mu}_0}
\newcommand{\phimub}[0]{\Phi^{\mathcal{B},\mu}_0}
\newcommand{\xhimub}[0]{X^{\mathcal{B},\mu}_0}
\newcommand{\psimub}[0]{\Psi^{\mathcal{B},\mu}_0}
\newcommand{\phiimub}[0]{\Phi^{\mathcal{B},\mu}_i}
\newcommand{\basis}[0]{\mathcal{B}}
\newcommand{\energyex}[0]{E^{\text{ex}}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\identity}{\mathds{1}}
\newcommand{\mur}[1]{\mu({\bf r_{#1}})}
\newcommand{\muueg}{\langle \mu_{\text{UEG}}\rangle}
\newcommand{\mursc}{\langle \mu_{r_{s,c}}\rangle}
\newcommand{\mursclda}{\langle \mu_{r_{s,c}^{\text{UEG}}}\rangle}

@@

\section{Brief overview of mathematical properties of similarity transformed Hamiltonian}
\subsection{Definition of the similarity transformation}
Assume an \textit{Hermitian operator} $H$ whose spectral representation reads 
\begin{equation}
 \label{eq:h_spectr}
 H  = \sum_n E_n \ket{\Psi_n}\bra{\Psi_n}
\end{equation}
with $\ket{\Psi_n}$ being its orthonormal eigenvectors 
\begin{equation}
 \bra{\Psi_n} \Psi_m \rangle = \delta_{mn},
\end{equation}
and $E_n \in \R$ its real eigenvalues.

Let us consider now an \textit{Hermitian} and \textit{invertible} operator $U$, \textit{i.e.} an operator $U$ such that there exists $U^{-1}$ such that 
\begin{equation}
 \label{eq:u_def_1}
 U^{-1} U = U \, U^{-1} = 1,
\end{equation}
and that 
\begin{equation}
 \begin{aligned}
 \label{eq:u_def_2}
 & U^\dagger = U \\
 & \big(U^{-1}\big)^{\dagger} = U^{-1}.
 \end{aligned}
\end{equation}
Then, the similarity transformed Hamiltonian $\tilde{H}$ is defined as
\begin{equation}
 \label{eq:def_htilde}
 \tilde{H} = U^{-1}\, H \, U.
\end{equation}
\subsection{Eigenvalues ans eigenvectors of $\tilde{H}$}
\subsubsection{Definition of left and right eigenvectors}
Regarding the properties of $\tilde{H}$, a first remark is that it is not Hermitian anymore
\begin{equation}
 \begin{aligned}
 \tilde{H}^\dagger & = \bigg(  U^{-1} \, H \,U \bigg)^\dagger \\
                   & = U\, H \, U^{-1}  \\
                   & \ne \tilde{H}.
 \end{aligned}
\end{equation}
Even though $\tilde{H}$ is not hermitian, it conserves the same eigenvalues than the original Hamiltonian $H$. 
The main difference with the original Hamiltonian are its eigenvectors: because $\tilde{H}$ is not Hermitian anymore, it possesses \textit{right} eigenvectors and \textit{left} eigenvectors. 
The right eigenvectors of a given operator $A$ are defined by
\begin{equation}
 A \ket{\Phi_n} = A_n \ket{\Phi_n}, 
\end{equation}
whereas the left eigenvectors of such operator are defined through
\begin{equation}
 A^{\dagger} \ket{X_n} = A_n \ket{X_n}.
\end{equation}
Therefore, if $A$ is hermitian, $A = A^{\dagger}$ and therefore $\ket{\Phi_n} = \ket{X_n} $, but for a non hermitian operator, the left and right eigenvectors do not coincide.

To find the right and left eigenvectors of $\tilde{H}$, one can insert the definition of $\tilde{H}$ (see Eq. \eqref{eq:def_htilde}) in Eq. \eqref{eq:h_spectr}
 \label{eq:tilde_1}
 \tilde{H} & = \sum_n E_n \, U^{-1} \, \ket{\Psi_n} \bra{\Psi_n}U.
We can define the right eigenvectors $\ket{\Phi_n}$ as
 \label{eq:def_right}
 \ket{\Phi_n} = U^{-1} \ket{\Psi_n},
 because  from Eq. \eqref{eq:tilde_1}, we notice that the $\ket{\Phi_n}$ fulfil the right eigenvector relationship
 \tilde{H} \ket{\Phi_n} & = \sum_m E_m \,U^{-1} \, \ket{\Psi_m} \bra{\Psi_m} U\ket{\Phi_n} \\
                        & = \sum_m E_m \,U^{-1} \, \ket{\Psi_m} \bra{\Psi_m} U \bigg( U^{-1} \ket{\Psi_n} \bigg) \\
                        & = \sum_m E_m \,U^{-1} \, \ket{\Psi_m} \bra{\Psi_m} \Psi_n \rangle \\
                        & = E_n          U^{-1} \ket{\Psi_n} \\
                        & = E_n  \ket{\Phi_n} \\
In a similar way, we can define the left eigenvectors $\ket{X_n}$ as 
 \label{eq:def_left}
 \ket{X_n} = U \ket{\Psi_n},
because they fulfil the left eigenvalue equation
 \tilde{H}^{\dagger} \ket{X_n}  = E_n \, \ket{X_n}. 
\subsubsection{Biorthogonality, spectral representation and loose of the variational principle}
The left and right eigenvectors fulfill the so-called \textit{biorthogonal} property, as 
from Eqs. \eqref{eq:def_right} and \eqref{eq:def_left} one obtains 
 \overlap{X_j}{\Phi_i} = \bigg(U \ket{\Psi_n} \bigg)^\dagger \bigg(U^{-1}\ket{\Psi_J}\bigg)
which, according to the Hermitian property of $U$ (see Eq.\eqref{eq:u_def_2}) simply becomes 
 \overlap{X_j}{\Phi_i} &= \matelem{\Psi_j}{U\,U^{-1}}{\Psi_i} \\
                       &= \overlap{\Psi_j}{\Psi_i} \\
                       &= \delta_{ji}.
Nevertheless, the right and left eigenvectors among themselves are not orthonormal as 
  \overlap{\Phi_j}{\Phi_i} & = \matelem{\Psi_j}{U^{-1}\,U^{-1}}{\Psi_i} \\
                           & = \matelem{\Psi_j}{U^{-2}}{\Psi_i},  
  \overlap{X_j}{X_i} & = \matelem{\Psi_j}{U\,U}{\Psi_i} \\
                           & = \matelem{\Psi_j}{U^{2}}{\Psi_i}.  
Having obtained biorthonormal left and right eigenvectors, one can then obtain a spectral representation of the similarity transformed Hamiltonian 
 \tilde{H} = \sum_i E_i \ket{\Phi_i} \bra{X_i}, 
which is very useful to study the variational principle, an important aspect of quantum chemistry. 
Let us consider a wave function $\Psi$ developed on the eigenvectors of $H$
 \ket{\Psi} = \sum_i c_i \ket{\Psi_i}.
The expectation value of $H$ over $\Psi$ is then 
  \matelem{\Psi}{H}{\Psi} & = \sum_{i,j} c_i c_j \matelem{\Psi_j}{H}{\Psi_i} \\
                          & = \sum_{i,j} c_i c_j  E_i \overlap{\Psi_j}{\Psi_i} \\
                          & = \sum_{i} |c_i|^2 E_i \\
                          & \ge E_0
which gives the variational principle because of the orthogonality of eigenvectors of an \textit{Hermitian} operator. 
Now let us decompose the same wave function $\Psi$ on the right-eigenvectors of $\tilde{H}$ 
 \ket{\Psi} = \sum_i \tilde{c}_i \ket{\Phi_i},
and consider the expectation value of $\tilde{H}$ over $\Psi$
  \matelem{\Psi}{\tilde{H}}{\Psi} & = \sum_{i,j} \tilde{c}_i \tilde{c}_j \matelem{\Phi_j}{\tilde{H}}{\Phi_i} \\
                                  & = \sum_{i,j} \tilde{c}_i \tilde{c}_j E_i \overlap{\Phi_j}{\Phi_i} 
which has no reason to be an upper bound to $E_0$ as $\overlap{\Phi_j}{\Phi_i} \ne \delta_{i,j}$. 
Therefore, we see that we cannot apply straightforwardly the variational principle 
to similarity transformed Hamiltonians, because of the lack of orthonormality of the right eigenvectors.

\section{A new form of Jastrow factor for transcorrelated Hamiltonian}
\subsection{The similarity transformed Hamiltonian}
Let us write the Hamiltonian of the He atom in the ${\bf r}_{12} = \big|{\bf r}_1 - {\bf r}_2 \big|$ coordinate 
\begin{equation}
 H  = h_c + \frac{1}{r_{12}},
\end{equation}
where 
\begin{equation}
 \begin{aligned}
 \label{def:h_c}
 h_c = &-\frac{1}{2} \sum_{i=1}^2 \bigg(\deriv{}{r_i}{2} + \frac{2}{r_i} \deriv{}{r_i}{} + \frac{2 Z}{r_i}\bigg) \\
     &-\bigg( \deriv{}{r_{12}}{2} + \frac{2}{r_{12}} \deriv{}{r_{12}}{} \bigg) \\
     &-\bigg( \frac{\bd{r_1}}{r_1} \cdot \frac{\bd{r_{12}} }{r_{12}}  \deriv{}{r_1}{} + 
                \frac{\bd{r_2}}{r_2} \cdot \frac{\bd{r_{21}} }{r_{21}}  \deriv{}{r_2}{} \bigg),
 \end{aligned}
\end{equation}
\label{sec:he_j}
Now let us consider the Similarity transformed Hamiltonian $\tilde{H}[j]$
\begin{equation}
 \label{eq:ht_0}
 \begin{aligned}
 \tilde{H}[j]&= e^{-j(r_{12})} H e^{j(r_{12})},
 \end{aligned}
\end{equation}
where $j(r_{12})$ is a general Jastrow factor depending only on $r_{12}$. 
Therefore, the only new terms arising in $\tilde{H}[j]$ with respect to $H$ are those coming from the differential operator in $r_{12}$,
\begin{equation}
 \mathcal{T}[j] =  -e^{-j(r_{12})}\bigg( \deriv{}{r_{12}}{2} + \frac{2}{r_{12}} \deriv{}{r_{12}}{} \bigg)e^{j(r_{12})}.  
\end{equation}
Let us compute the action of $\mathcal{T}[j]$ on a general function $\phi(\br{}_1,\br{}_2)$ 
\begin{equation}
 \begin{aligned}
 \mathcal{T}[j]\phi(\br{}_1,\br{}_2) & =- e^{-j(r_{12})} \bigg[ \deriv{}{r_{12}}{2} + \frac{2}{r_{12}} \deriv{}{r_{12}}{}\bigg]  e^{j(r_{12})} \phi(\br{}_1,\br{}_2) \\ 
                            &=  -\deriv{\phi(\br{}_1,\br{}_2)}{r_{12}}{2} - \frac{2}{r_{12}} \deriv{\phi(\br{}_1,\br{}_2)}{r_{12}}{} \\
  &- 2 \deriv{j(r_{12})}{r_{12}}{} \deriv{\phi(\br{}_1,\br{}_2)}{r_{12}}{} \\ 
  &- \bigg[ \frac{2}{r_{12}} \deriv{j(r_{12})}{r_{12}}{}  + \deriv{j(r_{12})}{r_{12}}{2} + \bigg( \deriv{j(r_{12})}{r_{12}}{} \bigg)^2\bigg] \phi(\br{}_1,\br{}_2). 
 \end{aligned}
\end{equation}
Now, by defining the following operators 
\begin{equation}
 \label{eq:def_tt}
 \tilde{t}[j] = -2 \deriv{j(r_{12})}{r_{12}}{} \deriv{}{r_{12}}{},
\end{equation}
\begin{equation}
 \label{eq:def_wt}
 \tilde{W}[j] = -\frac{2}{r_{12}} \deriv{j(r_{12})}{r_{12}}{}  , 
\end{equation}
\begin{equation}
 \label{eq:def_wt}
 \tilde{w}[j] = -\deriv{j(r_{12})}{r_{12}}{2} - \bigg( \deriv{j(r_{12})}{r_{12}}{} \bigg)^2. 
\end{equation}
one can write the Similarity transformed Hamiltonian  as
\begin{equation}
 \label{eq:ht_phi}
 \tilde{H}[j] = h_c + \frac{1}{r_{12}}  + \tilde{W}[j] + \tilde{t}[j] + \tilde{w}[j] ,
\end{equation}
ore more explicitly 
\begin{equation}
 \tilde{H}[j] = h_c + \frac{1}{r_{12}}\bigg( 1 - 2 \deriv{j(r_{12})}{r_{12}}{}\bigg) + \tilde{t}[j] + \tilde{w}[j] .
\end{equation}
$\tilde{H}[j]$, because of the first-order differential operator in $\tilde{t}[j]$, is then non-hermitian,   
and the effective potential $ \frac{1}{r_{12}}  + \tilde{W}[j] + \tilde{w}[j]$ in $\tilde{H}[j]$ only depends on the shape of $j(r_{12})$.

\subsection{$r_{12} \rightarrow 0$ limits for different forms of Hamiltonians}
\label{sec:cusp}
Having established the general form of the transcorrelated Hamiltonian $\tilde{H}[j]$, in the case of the helium atom, one can then study the behaviour of the latter when $r_{12}\rightarrow 0$ and compare it to two other types of Hamiltonian: the usual physical Hamiltonian and that entering the RS-DFT framework. 
The analysis of the $r_{12}\rightarrow 0$ behaviour leads to the different conditions that the eigenvectors of such operators must fulfill. For the sake of simplicity of the notations we focus on the ground state of these operators but the arguments are general for any bounded eigenvectors.

@@

Regarding now the similarity transformed Hamiltonians, the exact ground state eigenvector must also satisfy an eigenvalue equation in real-space,
\begin{equation}
 \label{eq:hpi0_ex}
 \tilde{H}[j] \phiex(\br{}_1,\br{}_2) = E_0 \phiex(\br{}_1,\br{}_2) \quad \forall (\br{}_1,\br{}_2).
\end{equation}
Similarly to Eq. \eqref{eq:cusp_0}, when looking at $r_{12}\approx 0$, the terms multiplying $\frac{1}{r_{12}}$ in $\tilde{H}[j]\phiex(\br{}_1,\br{}_2)$ must remain finite, 
which involves also the term $\tilde{W}[j]$ in $\tilde{H}[j]$ (see Eq.\eqref{eq:def_wt} and \eqref{eq:ht_phi}. Therefore, imposing the finiteness of the limit $r_{12}\rightarrow 0$ of Eq.\eqref{eq:hpi0_ex} translates into
\begin{equation}
 \label{eq:cusp_phi_0}
 \lim_{r_{12}\rightarrow 0}\bigg( \frac{2}{r_{12}} \deriv{}{r_{12}}{} -\frac{1}{r_{12}}\bigg[ 1 - 2 \deriv{j(r_{12})}{r_{12}}{}\bigg]\bigg) \phiex(\bd{r_1},\bd{r_2})  = c,\quad c<\infty.
\end{equation}
Then, the term $\frac{1}{r_{12}}\bigg[ 1 - 2\deriv{j(r_{12})}{r_{12}}{}\bigg]$ in Eq.\eqref{eq:cusp_phi_0} looks like an effective electron electron interaction modified by the presence of the Jastrow factor in $\tilde{H}[j]$. 
Therefore, as long as one imposes that 
\begin{equation}
 \label{eq:cusp_phi_1}
  \deriv{j(r_{12})}{r_{12}}{}\bigg|_{r_{12}=0} = \frac{1}{2},
\end{equation}
which is nothing but the cusp condition for the Jastrow factor $j(r_{12})$, the scalar term proportional to $\frac{1}{r_{12}}$ in Eq.\eqref{eq:cusp_phi_0} vanishes at $r_{12}=0$, and then obtain a non-divergent effective electron-electron interaction. 
Within the condition of Eq.\eqref{eq:cusp_phi_1}, the general condition of Eq.\eqref{eq:cusp_phi_0} becomes 
\begin{equation}
 \label{eq:cusp_phi_2}
 \deriv{\phiex(\bd{r_1},\bd{r_2})}{r_{12}}{}\Bigr|_{r_{12}=0} = 0, 
\end{equation}
which implies that as long as the Jastrow factor contains the cusp conditions, the eigenfunction of $\tilde{H}[j]$ are cuspless. 
For instance, a simple Jastrow factor of the form $j(r_{12}) = \frac{1}{2} r_{12}$ release $\phiex(\br{}_1,\br{}_2)$ from 
the constraint of fulfilling the cusp condition. 
%Another way to phrase Eq.\eqref{eq:cusp_phi_1} and Eq.\eqref{eq:cusp_phi_2} is that because of the Jastrow factor, the similarity transformed Hamiltonian exhibits an effective non-divergent electron-electron interaction, which does not impose any cusp condition to the eigenfunctions of $\tilde{H}[j]$.

@@

\subsection{Explicit form of  $\tilde{H}[\mu]$ and some of its properties}
\subsubsection{Explicit form of $\tilde{H}[\mu]$ }
Having established the explicit form of $j(r_{12};\mu)$ and some of its properties, one can now 
look at the explicit form of the transcorrelated Hamiltonian obtained with $j(r_{12};\mu)$,  
which, by injecting the form of \eqref{eq:def_j} in Eq.\eqref{eq:ht_phi}, reads 
 \label{eq:exp_ht_mu}
   \tilde{H}[\mu] = & h_c + \frac{\text{erf}(\mu r_{12})}{r_{12}} + \frac{\mu}{\sqrt{\pi}} e^{-\big(\mu r_{12} \big)^2} - \frac{\bigg(1 -     \text{erf}(\mu r_{12}) \bigg)^2}{4} \\
& + \bigg( \text{erf}(\mu r_{12}) - 1\bigg) \deriv{}{r_{12}}{}.
From Eq.\eqref{eq:exp_ht_mu}, one can define an effective scalar electron-electron interaction $\tilde{\mathcal{W}}_{ee}(r_{12}) $ 
 \label{eq:deff_w_new}
 \tilde{\mathcal{W}}_{ee}(r_{12},\mu) = \frac{\text{erf}(\mu r_{12})}{r_{12}} + \frac{\mu}{\sqrt{\pi}} e^{-\big(\mu r_{12} \big)^2} - \frac{\bigg(1 -     \text{erf}(\mu r_{12}) \bigg)^2}{4},
whose shape is illustrated in Fig\ref{fig_wee_mu}. 

>>>>

\definecolor{darkgreen}{HTML}{009900}
\usepackage[normalem]{ulem}
\newcommand{\sphi}[1]{\hat{{\bf S}}_{#1}}
\newcommand{\overlap}[2]{\langle #1 | #2 \rangle}
\newcommand{\matelem}[3]{\langle #1 | #2 | #3 \rangle}
\newcommand{\deriv}[3]{\frac{\partial^{#3} #1}{\partial {#2}^{#3}}}
\newcommand{\bd}[1]{{\bf {#1}}}
\newcommand{\br}[0]{{\bf {r}}}
\newcommand{\bri}[1]{{\bf r}_{#1}}
\newcommand{\bs}[0]{{\bf {s}}}
\newcommand{\dr}[1]{\text{d}{\bf r_{#1}}}
\newcommand{\psiex}[0]{\Psi^{\text{ex}}_0}
\newcommand{\phiex}[0]{\Phi^{\text{ex}}_0}
\newcommand{\phimu}[0]{\Phi^{\text{ex},\mu}_0}
\newcommand{\phimub}[0]{\Phi^{\mathcal{B},\mu}_0}
\newcommand{\xhimub}[0]{X^{\mathcal{B},\mu}_0}
\newcommand{\psimub}[0]{\Psi^{\mathcal{B},\mu}_0}
\newcommand{\phiimub}[0]{\Phi^{\mathcal{B},\mu}_i}
\newcommand{\basis}[0]{\mathcal{B}}
\newcommand{\energyex}[0]{E^{\text{ex}}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\identity}{\mathds{1}}
\newcommand{\mur}[1]{\mu({\bf r_{#1}})}
\newcommand{\muueg}{\mu_{\text{UEG}}}
\newcommand{\muuegav}{\langle \mu_{\text{UEG}}\rangle}
\newcommand{\mursc}{ \mu_{r_{s,c}}}
\newcommand{\murscav}{\langle \mu_{r_{s,c}}\rangle}
\newcommand{\mursclda}{\langle \mu_{r_{s,c}^{\text{UEG}}}\rangle}

@@

The paper is organized as follows. 
In the first section, we derive the analytical form of a new Jastrow factor and the corresponding transcorrelated Hamiltonian. 
The, we perform a detailed numerical study of the behaviour of the solution of $\tilde{H}[\mu]$ on the helium atom: 
in the first part we analyze the speed of convergence of the ground state energy with respect to the basis set and the parameter $\mu$, 
then we study the behaviour of the right eigenvectors in real space and compare it with the exact ground state wave function. 
We show that the rapid convergence of the total energies coincides with a rapid convergence of the right eigenvectors in real space. 
Based on such encouraging results, we study the helium isoelectronic series from Li$^+$ to Ne$^{8+}$ and compare different schemes to obtain 
a value of $\mu$ which depends on the system and still provides a fast convergence of the total energies. 
Eventually, we study the H$_2$ molecule and compare different regimes of correlation.

\section{A new form of Jastrow factor for transcorrelated Hamiltonian}
This section is dedicated to the derivation of a new form of Jastrow factor mimicking at short range the correlation effects of RS-DFT. 
For the sake of simplicity and to remain as close as possible to the physical picture of the electronic correlation problem, 
we derive the equations of the transcorrelated Hamiltonian for the helium atom for a general Jastrow factor depending on the inter electronic distance.  
Thanks to the simplicity of the equations, we focus on the behaviour at small $r_{12}$ of different flavours of Hamiltonians 
and the corresponding discontinuity in their eigenvectors. 
We browse three different types of Hamiltonians: the usual physical Hamiltonian, the transcorrelated Hamiltonian together with the effective Hamiltonian used in RS-DFT.
Then, we derive the general analytical form of the corresponding new transcorrelated Hamiltonians $\tilde{H}[\mu]$, and its dependence as function of the parameter $\mu$.

\subsection{A simple physical picture: the case of the helium atom}
\subsubsection{The transcorrelated Hamiltonian for helium atom}
Let us write the Hamiltonian of the He atom in the ${\bf r}_{12} = \big|{\bf r}_1 - {\bf r}_2 \big|$ coordinate 
\begin{equation}
 H  = h_c + \frac{1}{r_{12}},
\end{equation}
where 
\begin{equation}
 \begin{aligned}
 \label{def:h_c}
 h_c = &-\frac{1}{2} \sum_{i=1}^2 \bigg(\deriv{}{r_i}{2} + \frac{2}{r_i} \deriv{}{r_i}{} + \frac{2 Z}{r_i}\bigg) \\
     &-\bigg( \deriv{}{r_{12}}{2} + \frac{2}{r_{12}} \deriv{}{r_{12}}{} \bigg) \\
     &-\bigg( \frac{\bd{r_1}}{r_1} \cdot \frac{\bd{r_{12}} }{r_{12}}  \deriv{}{r_1}{} + 
                \frac{\bd{r_2}}{r_2} \cdot \frac{\bd{r_{21}} }{r_{21}}  \deriv{}{r_2}{} \bigg),
 \end{aligned}
\end{equation}
\label{sec:he_j}
Now let us consider the Similarity transformed Hamiltonian $\tilde{H}[j]$
\begin{equation}
 \label{eq:ht_0}
 \begin{aligned}
 \tilde{H}[j]&= e^{-j(r_{12})} H e^{j(r_{12})},
 \end{aligned}
\end{equation}
where $j(r_{12})$ is a general Jastrow factor depending only on $r_{12}$. 
Therefore, the only new terms arising in $\tilde{H}[j]$ with respect to $H$ are those coming from the differential operator in $r_{12}$,
\begin{equation}
 \mathcal{T}[j] =  -e^{-j(r_{12})}\bigg( \deriv{}{r_{12}}{2} + \frac{2}{r_{12}} \deriv{}{r_{12}}{} \bigg)e^{j(r_{12})}.  
\end{equation}
Let us compute the action of $\mathcal{T}[j]$ on a general function $\phi(\br{}_1,\br{}_2)$ 
\begin{equation}
 \begin{aligned}
 \mathcal{T}[j]\phi(\br{}_1,\br{}_2) & =- e^{-j(r_{12})} \bigg[ \deriv{}{r_{12}}{2} + \frac{2}{r_{12}} \deriv{}{r_{12}}{}\bigg]  e^{j(r_{12})} \phi(\br{}_1,\br{}_2) \\ 
                            &=  -\deriv{\phi(\br{}_1,\br{}_2)}{r_{12}}{2} - \frac{2}{r_{12}} \deriv{\phi(\br{}_1,\br{}_2)}{r_{12}}{} \\
  &- 2 \deriv{j(r_{12})}{r_{12}}{} \deriv{\phi(\br{}_1,\br{}_2)}{r_{12}}{} \\ 
  &- \bigg[ \frac{2}{r_{12}} \deriv{j(r_{12})}{r_{12}}{}  + \deriv{j(r_{12})}{r_{12}}{2} + \bigg( \deriv{j(r_{12})}{r_{12}}{} \bigg)^2\bigg] \phi(\br{}_1,\br{}_2). 
 \end{aligned}
\end{equation}
Now, by defining the following operators 
\begin{equation}
 \label{eq:def_tt}
 \tilde{t}[j] = -2 \deriv{j(r_{12})}{r_{12}}{} \deriv{}{r_{12}}{},
\end{equation}
\begin{equation}
 \label{eq:def_wt}
 \tilde{W}[j] = -\frac{2}{r_{12}} \deriv{j(r_{12})}{r_{12}}{}  , 
\end{equation}
\begin{equation}
 \label{eq:def_wt}
 \tilde{w}[j] = -\deriv{j(r_{12})}{r_{12}}{2} - \bigg( \deriv{j(r_{12})}{r_{12}}{} \bigg)^2. 
\end{equation}
one can write the Similarity transformed Hamiltonian  as
\begin{equation}
 \label{eq:ht_phi}
 \tilde{H}[j] = h_c + \frac{1}{r_{12}}  + \tilde{W}[j] + \tilde{t}[j] + \tilde{w}[j] ,
\end{equation}
ore more explicitly 
\begin{equation}
 \tilde{H}[j] = h_c + \frac{1}{r_{12}}\bigg( 1 - 2 \deriv{j(r_{12})}{r_{12}}{}\bigg) + \tilde{t}[j] + \tilde{w}[j] .
\end{equation}
$\tilde{H}[j]$, because of the first-order differential operator in $\tilde{t}[j]$, is then non-hermitian,   
and the effective potential $ \frac{1}{r_{12}}  + \tilde{W}[j] + \tilde{w}[j]$ in $\tilde{H}[j]$ only depends on the shape of $j(r_{12})$.

\subsubsection{$r_{12} \rightarrow 0$ limits for different forms of Hamiltonians}
\label{sec:cusp}
Having established the general form of the transcorrelated Hamiltonian $\tilde{H}[j]$, in the case of the helium atom, one can then study the behaviour of the latter when $r_{12}\rightarrow 0$ and compare it to two other types of Hamiltonian: the usual physical Hamiltonian and that entering the RS-DFT framework. 
The analysis of the $r_{12}\rightarrow 0$ behaviour leads to the different conditions that the eigenvectors of such operators must fulfill. For the sake of simplicity of the notations we focus on the ground state of these operators but the arguments are general for any bounded eigenvectors.

@@

Regarding now the similarity transformed Hamiltonians, the exact ground state eigenvector must also satisfy an eigenvalue equation in real-space,
\begin{equation}
 \label{eq:hpi0_ex}
 \tilde{H}[j] \phiex(\br{}_1,\br{}_2) = E_0 \phiex(\br{}_1,\br{}_2) \quad \forall (\br{}_1,\br{}_2).
\end{equation}
Similarly to Eq. \eqref{eq:cusp_0}, when looking at $r_{12}\approx 0$, the terms multiplying $\frac{1}{r_{12}}$ in $\tilde{H}[j]\phiex(\br{}_1,\br{}_2)$ must remain finite, 
which involves also the term $\tilde{W}[j]$ in $\tilde{H}[j]$ (see Eq.\eqref{eq:def_wt} and \eqref{eq:ht_phi}. Therefore, imposing the finiteness of the limit $r_{12}\rightarrow 0$ of Eq.\eqref{eq:hpi0_ex} translates into
\begin{equation}
 \label{eq:cusp_phi_0}
 \lim_{r_{12}\rightarrow 0}\bigg( \frac{2}{r_{12}} \deriv{}{r_{12}}{} -\frac{1}{r_{12}}\bigg[ 1 - 2 \deriv{j(r_{12})}{r_{12}}{}\bigg]\bigg) \phiex(\bd{r_1},\bd{r_2})  = c,\quad c<\infty.
\end{equation}
Then, the term $\frac{1}{r_{12}}\bigg[ 1 - 2\deriv{j(r_{12})}{r_{12}}{}\bigg]$ in Eq.\eqref{eq:cusp_phi_0} looks like an effective electron electron interaction modified by the presence of the Jastrow factor in $\tilde{H}[j]$. 
Therefore, as long as one imposes that 
\begin{equation}
 \label{eq:cusp_phi_1}
  \deriv{j(r_{12})}{r_{12}}{}\bigg|_{r_{12}=0} = \frac{1}{2},
\end{equation}
which is nothing but the cusp condition for the Jastrow factor $j(r_{12})$, the scalar term proportional to $\frac{1}{r_{12}}$ in Eq.\eqref{eq:cusp_phi_0} vanishes at $r_{12}=0$, and then obtain a non-divergent effective electron-electron interaction. 
Within the condition of Eq.\eqref{eq:cusp_phi_1}, the general condition of Eq.\eqref{eq:cusp_phi_0} becomes 
\begin{equation}
 \label{eq:cusp_phi_2}
 \deriv{\phiex(\bd{r_1},\bd{r_2})}{r_{12}}{}\Bigr|_{r_{12}=0} = 0, 
\end{equation}
which implies that as long as the Jastrow factor contains the cusp conditions, the eigenfunction of $\tilde{H}[j]$ are cuspless. 
For instance, a simple Jastrow factor of the form $j(r_{12}) = \frac{1}{2} r_{12}$ release $\phiex(\br{}_1,\br{}_2)$ from 
the constraint of fulfilling the cusp condition.

@@

\subsection{The general analytical form of  $\tilde{H}[\mu]$ and some of its properties}
\subsubsection{Explicit form of $\tilde{H}[\mu]$ }
Having established the analytical expression of $j(r_{12};\mu)$, one can derive the general expression of the associated transcorrelated Hamiltonian.  
According to Eq. (2) of Ref. \onlinecite{CohLuoGutDobTewAla-JCP-19}, the similarity transformed Hamiltonian can be written as 
\begin{equation}
 \label{ht_def_g}
 e^{-\hat{\tau}} \hat{H} e^{\hat{\tau}} = H + \big[ H,\hat{\tau} \big] + \frac{1}{2}\bigg[ \big[H,\hat{\tau}\big],\hat{\tau}\bigg]
\end{equation}
where $\hat{\hat{\tau}} = \sum_{i>j}u(\br{}_i,\br{}_j)$ and $\hat{H} = \sum_i -\frac{1}{2} \nabla^2_i + v(\br{}_i) + \sum_{i>j} \frac{1}{r_{ij}}$. 
This leads to the following similarity transformed Hamiltonian 
\begin{equation}
 \begin{aligned}
 \label{ht_def_gu}
 \tilde{H} & = H - \sum_i \bigg( \frac{1}{2} \nabla_i^2 \hat{\tau} + \big(\nabla_i \hat{\tau} \big) + \frac{1}{2} \big(\nabla_i \hat{\tau} \big)  \bigg) \\
           & = H - \sum_{i<j} \hat{K}(\bri{i},\bri{j}) - \sum_{i<j<k} \hat{L}(\bri{i},\bri{j},\bri{k}),
 \end{aligned}
\end{equation}
where the effective two- and three-body operators $\hat{K}(\bri{i},\bri{j})$ and $\hat{L}(\bri{i},\bri{j},\bri{k})$ are defined as
\begin{equation}
 \begin{aligned}
 \label{def_k}
  \hat{K}(\bri{1},\bri{2}) ) = \frac{1}{2} \bigg( &\nabla_1^2 u(\bri{1},\bri{2}) + \nabla_2^2u(\bri{1},\bri{2}) \\
                                               + &\big(\nabla_1 u(\bri{1},\bri{2}) \big) ^2 + \big(\nabla_1 u(\bri{1},\bri{2}) \big) ^2 \bigg) \\
                                               + &\nabla_1 u(\bri{1},\bri{2}) \cdot \nabla_2 + \nabla_2 u(\bri{1},\bri{2}) \cdot \nabla_1  \\
 \end{aligned}
\end{equation}
\begin{equation}
 \begin{aligned}
 \label{def_k}
  \hat{L}(\bri{1},\bri{2},\bri{3}) ) = & \nabla_1 u(\bri{1},\bri{2}) \cdot \nabla_1 u(\bri{1},\bri{3}) + \nabla_2 u(\bri{2},\bri{1}) \cdot \nabla_2 u(\bri{2},\bri{3})  \\
                                     + & \nabla_3 u(\bri{3},\bri{1}) \cdot \nabla_3 u(\bri{3},\bri{2}).
 \end{aligned}
\end{equation}
Therefore, to obtain the general formulation of the $\tilde{H}[\mu]$ one needs to compute the actual form of the operators $\hat{K}(\bri{1},\bri{2})$ and $\hat{L}(\bri{1},\bri{2},\bri{3})$ when $u(\bri{1},\bri{2}) = j(r_{12},\mu)$.
